### 2020-02-16 WorkingCode/Project1_RM_Objective_1.2.16.20.Rmd
Add files via upload

---
title: "Project1_RM_Objective_1"
Date: 2/10/2020
Author: Reagan Meagher
---

----------
  SETUP
----------

```{r setup and clear envirornment, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
rm(list = ls()) #Clear envirornment code
options(scipen=999) #No exponential formatting of numbers
options(max.print=1000000)
```

```{r libraries}
library(tidyverse)
library(caret)
library(mice)
library(skimr)
library(corrplot)
library(car)
library(ISLR)
library(ggplot2)
library(gridExtra)
library(SamplingStrata)
library(rbin)
library(leaps)
library(dplyr)
library(ggplot2)
library(geosphere)
library(broom)
library(plyr)
```

----------
   EDA
----------

```{r EDA, create linear-linear data set}
nycraw <- read.csv("https://raw.githubusercontent.com/JaclynCoate/6372_Project_1/master/AB_NYC_2019.csv", header = TRUE, strip.white=TRUE)
nyc <- nycraw

#Creating a new, tsquare_distance (distance to Times Square in miles)
for (i in 1:nrow(nyc)) {
  n <- (distHaversine(c(nyc$latitude[i], -nyc$longitude[i]), c(40.7580, 73.9855)) / 1609.344)
  nyc$tsquare_distance[i] <- n
}

#Dropping logical irrelevant variables: "id", "name", "host_id", "host_name", "last_review", "latitude", "longitude", "neighborhood"
nyc <- select(nyc, -c("id", "name", "host_id", "host_name", "last_review", "latitude", "longitude", "neighbourhood"))

#### Zero Value Variable Check
#- Checking on dependent variable range to make sure if there are zero's to remove. It would not be free to stay in NYC.
#- Checking on independent variables that have '0' as a value. Seeing that "availability_365" would mean the AirBnB isn't available at all, we are dropping those from our data set.
nyc <- nyc[!(nyc$price==0),]
nyc <- nyc[!(nyc$availability_365==0),]

#Checking for NAs
md.pattern(nyc)
#Drop NAs that are present
nyc <- na.omit(nyc)
#Confirming NA drop
nrow(nyc)

#Storing categorical variables as factors
skim(nyc)

#Removing reviews_per_month due to high correlation of is and number_of_reviews
nyc <- select(nyc, -c("reviews_per_month"))

#Generate a summary of the current data set
#summary(nyc)

nyc <- filter(nyc, neighbourhood_group != "Staten Island")

#Restricting data set to prices between $25 and $400 a night (affordable market)
nyc <- filter(nyc, price >= 25 & price <= 400)

#Removing outliers of minimum nights stay
nyc <- nyc[!(nyc$minimum_nights > 365),]
#invisible(view(nyc))

#Creating linear-linear data set
linear.nyc <- nyc

#Linearity check of linear-linear data set
#pairs(nyc,col=nyc$neighbourhood_group) #Color by neighborhood
```

```{r create log-linear data set}
log.nyc <- nyc %>% mutate(lprice=log(price))
log.nyc <- select(log.nyc, -c("price"))
invisible(log.nyc)

#Linearity check of log-linear data set
#pairs(log.nyc,col=log.nyc$neighbourhood_group)
```

```{r create log-log data set}
log.log.nyc <- log.nyc %>% mutate(lreviews=log(number_of_reviews))
log.log.nyc <- log.log.nyc %>% mutate(lnights=log(minimum_nights))
log.log.nyc <- log.log.nyc %>% mutate(llistings=log(calculated_host_listings_count))
log.log.nyc <- log.log.nyc %>% mutate(lavailablility=log(availability_365))
log.log.nyc <- log.log.nyc %>% mutate(ltsqr=log(tsquare_distance))
invisible(log.log.nyc)

log.log.nyc <- select(log.log.nyc, -c("minimum_nights", "number_of_reviews", "calculated_host_listings_count", "availability_365", "tsquare_distance"))
# Checking for -inf logged results
invisible(log.log.nyc)
# Drop -inf log reults in lavailability
log.log.nyc<-log.log.nyc[!(log.log.nyc$lavailablility=="-Inf"),]
invisible(log.log.nyc)

#Linearity Check of log-log data set
#pairs(log.indep.nyc,col=log.indep.nyc$neighbourhood_group) #Color by neighborhood
```

```{r hold predict function}
#Really handy predict function
predict.regsubsets =function (object , newdata ,id ,...){
  form=as.formula (object$call [[2]])
  mat=model.matrix(form ,newdata )
  coefi=coef(object ,id=id)
  xvars=names(coefi)
  mat[,xvars]%*%coefi
}
```

----------
 MODELING
----------

TEST/TRAIN CREATION
```{r create 70/30 train/test splits for each data source}
set.seed(1234)

#nyc
index<-sample(1:dim(linear.nyc)[1],nrow(linear.nyc)*.7,replace=F)
linear.train<-linear.nyc[index,]
linear.test<-linear.nyc[-index,]

#log.nyc
index<-sample(1:dim(log.nyc)[1],nrow(log.nyc)*.7,replace=F)
log.train<-log.nyc[index,]
log.test<-log.nyc[-index,]
```

INTUITIVE LINEAR - LINEAR
```{r intuitive model building - linear}
#LINEAR - LINEAR
linear.intuitive.train.model <- lm(price ~ minimum_nights + number_of_reviews + tsquare_distance + neighbourhood_group + room_type + neighbourhood_group:room_type, data = linear.train)
linear.intuitive.predictions <- linear.intuitive.train.model %>% predict(linear.test)
print("")
print("Test Adjusted R Squared: ")
print(R2(linear.intuitive.predictions, linear.test$price))
print("Test Root Mean Square Error: ")
print(RMSE(linear.intuitive.predictions, linear.test$price))
print("Test Mean Absolute Error: ")
print(MAE(linear.intuitive.predictions, linear.test$price))
print("Prediction Error Rate: ")
RMSE(linear.intuitive.predictions, linear.test$price)/mean(linear.test$price)

linear.intuitive.final.model <- lm(price ~ minimum_nights + number_of_reviews + tsquare_distance + neighbourhood_group + room_type + neighbourhood_group:room_type, data = linear.nyc)
print("Full Model Adjusted R Squared: ")
print(summary(linear.intuitive.final.model)$adj.r.squared)
print("Full Model AIC: ")
print(AIC(linear.intuitive.final.model))
print("Full Model BIC: ")
print(BIC(linear.intuitive.final.model))
print("Full Model RSS: ")
print(sum(resid(linear.intuitive.final.model)^2))
summary(linear.intuitive.final.model)
par(mfrow=c(2,2))
plot(linear.intuitive.final.model)
vif(linear.intuitive.final.model)
confint(linear.intuitive.final.model)
```

INTUITIVE LOG - LINEAR
```{r intuitive model building - log}
#LOG - LINEAR
log.intuitive.test.model <- lm(lprice ~ minimum_nights + number_of_reviews + tsquare_distance + neighbourhood_group + room_type + neighbourhood_group:room_type, data = log.train)
log.intuitive.predictions <- log.intuitive.test.model %>% predict(log.test)
print("")
print("Test Adjusted R Squared: ")
print(R2(log.intuitive.predictions, log.test$lprice))
print("Test Root Mean Square Error: ")
print(RMSE(log.intuitive.predictions, log.test$lprice))
print("Test Mean Absolute Error: ")
print(MAE(log.intuitive.predictions, log.test$lprice))
print("Prediction Error Rate: ")
RMSE(log.intuitive.predictions, log.test$lprice)/mean(log.test$lprice)

log.intuitive.final.model <- lm(lprice ~ minimum_nights + number_of_reviews + tsquare_distance + neighbourhood_group + room_type + neighbourhood_group:room_type, data = log.nyc)
print("Full Model Adjusted R Squared: ")
print(summary(log.intuitive.final.model)$adj.r.squared)
print("Full Model AIC: ")
print(AIC(log.intuitive.final.model))
print("Full Model BIC: ")
print(BIC(log.intuitive.final.model))
print("Full Model RSS: ")
print(sum(resid(log.intuitive.final.model)^2))
summary(log.intuitive.final.model)
par(mfrow=c(2,2))
plot(log.intuitive.final.model)
vif(log.intuitive.final.model)
confint(log.intuitive.final.model)
```

------------------------------------------------------------------

FORWARD LINEAR
```{r forward selection model - linear}
#FORWARD LINEAR
linear.forward.select.model = regsubsets(price~.,data=linear.train,method="forward",nvmax=20)

par(mfrow=c(1,3))
fwd.bic <- summary(linear.forward.select.model)$bic
plot(1:length(summary(linear.forward.select.model)$bic), fwd.bic, type = "l", ylab = "BIC", xlab = "# Of Predictors")
fwd.index.bic <- which(fwd.bic == min(fwd.bic))
points(fwd.index.bic, fwd.bic[fwd.index.bic],col="red", pch=10)
print("")
print("Selected Number of Predictors From Forward Selection (BIC)")
print(fwd.index.bic)

fwd.adjr2 <- summary(linear.forward.select.model)$adjr2
plot(1:length(summary(linear.forward.select.model)$adjr2), fwd.adjr2, type = "l", ylab = "Adjusted R-squared", xlab = "# Of Predictors")
fwd.index.adjr2 <- which(fwd.adjr2 == max(fwd.adjr2))
points(fwd.index.adjr2, fwd.adjr2[fwd.index.adjr2],col="red", pch=10)
print("")
print("Selected Number of Predictors From Forward Selection (AR2)")
print(fwd.index.adjr2)

fwd.rss <- summary(linear.forward.select.model)$rss
plot(1:length(summary(linear.forward.select.model)$rss), fwd.rss, type = "l", ylab = "RSS", xlab = "# Of Predictors")
fwd.index.rss <- which(fwd.rss == min(fwd.rss))
points(fwd.index.rss, fwd.rss[fwd.index.rss],col="red", pch=10)
print("")
print("Selected Number of Predictors From Forward Selection (RSS)")
print(fwd.index.rss)

linear.forward.select.model2 = regsubsets(price~.,data=linear.train,method="forward",nvmax=(fwd.index.bic)+1)
coef(linear.forward.select.model2,fwd.index.bic)

linear.forward.train.model <-lm(price~neighbourhood_group + number_of_reviews + minimum_nights + calculated_host_listings_count + room_type + tsquare_distance,data=linear.train)
linear.forward.predictions <- linear.forward.train.model %>% predict(linear.test)
print("")
print("Test Adjusted R Squared: ")
print(R2(linear.forward.predictions, linear.test$price))
print("Test Root Mean Square Error: ")
print(RMSE(linear.forward.predictions, linear.test$price))
print("Test Mean Absolute Error: ")
print(MAE(linear.forward.predictions, linear.test$price))
print("Prediction Error Rate: ")
RMSE(linear.forward.predictions, linear.test$price)/mean(linear.test$price)

linear.forward.final.model <- lm(price~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance, data = linear.nyc)
print("Full Model Adjusted R Squared: ")
print(summary(linear.forward.final.model)$adj.r.squared)
print("Full Model AIC: ")
print(AIC(linear.forward.final.model))
print("Full Model BIC: ")
print(BIC(linear.forward.final.model))
print("Full Model RSS: ")
print(sum(resid(linear.forward.final.model)^2))
summary(linear.forward.final.model)
par(mfrow=c(2,2))
plot(linear.forward.final.model)
vif(linear.forward.final.model)
confint(linear.forward.final.model)
```

FORWARD LOG
```{r forward selection model - log}
#FORWARD LOG
log.forward.select.model = regsubsets(lprice~.,data=log.train,method="forward",nvmax=500)

par(mfrow=c(1,3))
fwd.bic <- summary(log.forward.select.model)$bic
plot(1:length(summary(log.forward.select.model)$bic), fwd.bic, type = "l", ylab = "BIC", xlab = "# Of Predictors")
fwd.index.bic <- which(fwd.bic == min(fwd.bic))
points(fwd.index.bic, fwd.bic[fwd.index.bic],col="red", pch=10)
print("")
print("Selected Number of Predictors From Forward Selection (BIC)")
print(fwd.index.bic)

fwd.adjr2 <- summary(log.forward.select.model)$adjr2
plot(1:length(summary(log.forward.select.model)$adjr2), fwd.adjr2, type = "l", ylab = "Adjusted R-squared", xlab = "# Of Predictors")
fwd.index.adjr2 <- which(fwd.adjr2 == max(fwd.adjr2))
points(fwd.index.adjr2, fwd.adjr2[fwd.index.adjr2],col="red", pch=10)
print("")
print("Selected Number of Predictors From Forward Selection (AR2)")
print(fwd.index.adjr2)

fwd.rss <- summary(log.forward.select.model)$rss
plot(1:length(summary(log.forward.select.model)$rss), fwd.rss, type = "l", ylab = "RSS", xlab = "# Of Predictors")
fwd.index.rss <- which(fwd.rss == min(fwd.rss))
points(fwd.index.rss, fwd.rss[fwd.index.rss],col="red", pch=10)
print("")
print("Selected Number of Predictors From Forward Selection (RSS)")
print(fwd.index.rss)

log.forward.select.model2 = regsubsets(lprice~.,data=log.train,method="forward",nvmax=(fwd.index.bic)+1)
coef(log.forward.select.model2,fwd.index.bic)

log.forward.train.model <-lm(lprice~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance,data=log.train)
log.forward.predictions <- log.forward.train.model %>% predict(log.test)
print("")
print("Test Adjusted R Squared: ")
print(R2(log.forward.predictions, log.test$lprice))
print("Test Root Mean Square Error: ")
print(RMSE(log.forward.predictions, log.test$lprice))
print("Test Mean Absolute Error: ")
print(MAE(log.forward.predictions, log.test$lprice))
print("Prediction Error Rate: ")
RMSE(log.forward.predictions, log.test$lprice)/mean(log.test$lprice)

log.forward.final.model <- lm(lprice~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance, data = log.nyc)
print("Full Model Adjusted R Squared: ")
print(summary(log.forward.final.model)$adj.r.squared)
print("Full Model AIC: ")
print(AIC(log.forward.final.model))
print("Full Model BIC: ")
print(BIC(log.forward.final.model))
print("Full Model RSS: ")
print(sum(resid(log.forward.final.model)^2))
summary(log.forward.final.model)
par(mfrow=c(2,2))
plot(log.forward.final.model)
vif(log.forward.final.model)
confint(log.forward.final.model)
```

BACKWARD LINEAR
```{r backward selection model - linear}
linear.backward.select.model = regsubsets(price~.,data=linear.train,method="backward",nvmax=500)

par(mfrow=c(1,3))
bkwrd.bic <- summary(linear.backward.select.model)$bic
plot(1:length(summary(linear.backward.select.model)$bic), bkwrd.bic, type = "l", ylab = "BIC", xlab = "# Of Predictors")
bkwrd.index.bic <- which(bkwrd.bic == min(bkwrd.bic))
points(bkwrd.index.bic, bkwrd.bic[bkwrd.index.bic],col="red", pch=10)
print("")
print("Selected Number of Predictors From backward Selection (BIC)")
print(bkwrd.index.bic)

bkwrd.adjr2 <- summary(linear.backward.select.model)$adjr2
plot(1:length(summary(linear.backward.select.model)$adjr2), bkwrd.adjr2, type = "l", ylab = "Adjusted R-squared", xlab = "# Of Predictors")
bkwrd.index.adjr2 <- which(bkwrd.adjr2 == max(bkwrd.adjr2))
points(bkwrd.index.adjr2, bkwrd.adjr2[bkwrd.index.adjr2],col="red", pch=10)
print("")
print("Selected Number of Predictors From backward Selection (AR2)")
print(bkwrd.index.adjr2)

bkwrd.rss <- summary(linear.backward.select.model)$rss
plot(1:length(summary(linear.backward.select.model)$rss), bkwrd.rss, type = "l", ylab = "RSS", xlab = "# Of Predictors")
bkwrd.index.rss <- which(bkwrd.rss == min(bkwrd.rss))
points(bkwrd.index.rss, bkwrd.rss[bkwrd.index.rss],col="red", pch=10)
print("")
print("Selected Number of Predictors From backward Selection (RSS)")
print(bkwrd.index.rss)

linear.backward.select.model2 = regsubsets(price~.,data=linear.train,method="backward",nvmax=(bkwrd.index.bic)+1)
coef(linear.backward.select.model2,bkwrd.index.bic)

linear.backward.train.model <-lm(price~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance,data=linear.train)
linear.backward.predictions <- linear.backward.train.model %>% predict(linear.test)
print("")
print("Test Adjusted R Squared: ")
print(R2(linear.backward.predictions, linear.test$price))
print("Test Root Mean Square Error: ")
print(RMSE(linear.backward.predictions, linear.test$price))
print("Test Mean Absolute Error: ")
print(MAE(linear.backward.predictions, linear.test$price))
print("Prediction Error Rate: ")
RMSE(linear.backward.predictions, linear.test$price)/mean(linear.test$price)

linear.backward.final.model <- lm(price~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance, data = linear.nyc)
print("Full Model Adjusted R Squared: ")
print(summary(linear.backward.final.model)$adj.r.squared)
print("Full Model AIC: ")
print(AIC(linear.backward.final.model))
print("Full Model BIC: ")
print(BIC(linear.backward.final.model))
print("Full Model RSS: ")
print(sum(resid(linear.backward.final.model)^2))
summary(linear.backward.final.model)
par(mfrow=c(2,2))
plot(linear.backward.final.model)
vif(linear.backward.final.model)
confint(linear.backward.final.model)
```

BACKWARD LOG
```{r backward selection model - log}
log.backward.select.model = regsubsets(lprice~.,data=log.train,method="backward",nvmax=500)

par(mfrow=c(1,3))
bkwrd.bic <- summary(log.backward.select.model)$bic
plot(1:length(summary(log.backward.select.model)$bic), bkwrd.bic, type = "l", ylab = "BIC", xlab = "# Of Predictors")
bkwrd.index.bic <- which(bkwrd.bic == min(bkwrd.bic))
points(bkwrd.index.bic, bkwrd.bic[bkwrd.index.bic],col="red", pch=10)
print("")
print("Selected Number of Predictors From backward Selection (BIC)")
print(bkwrd.index.bic)

bkwrd.adjr2 <- summary(log.backward.select.model)$adjr2
plot(1:length(summary(log.backward.select.model)$adjr2), bkwrd.adjr2, type = "l", ylab = "Adjusted R-squared", xlab = "# Of Predictors")
bkwrd.index.adjr2 <- which(bkwrd.adjr2 == max(bkwrd.adjr2))
points(bkwrd.index.adjr2, bkwrd.adjr2[bkwrd.index.adjr2],col="red", pch=10)
print("")
print("Selected Number of Predictors From backward Selection (AR2)")
print(bkwrd.index.adjr2)

bkwrd.rss <- summary(log.backward.select.model)$rss
plot(1:length(summary(log.backward.select.model)$rss), bkwrd.rss, type = "l", ylab = "RSS", xlab = "# Of Predictors")
bkwrd.index.rss <- which(bkwrd.rss == min(bkwrd.rss))
points(bkwrd.index.rss, bkwrd.rss[bkwrd.index.rss],col="red", pch=10)
print("")
print("Selected Number of Predictors From backward Selection (RSS)")
print(bkwrd.index.rss)

log.backward.select.model2 = regsubsets(lprice~.,data=log.train,method="backward",nvmax=(bkwrd.index.bic)+1)
coef(log.backward.select.model2,bkwrd.index.bic)

log.backward.train.model <-lm(lprice~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance,data=log.train)
log.backward.predictions <- log.backward.train.model %>% predict(log.test)
print("")
print("Test Adjusted R Squared: ")
print(R2(log.backward.predictions, log.test$lprice))
print("Test Root Mean Square Error: ")
print(RMSE(log.backward.predictions, log.test$lprice))
print("Test Mean Absolute Error: ")
print(MAE(log.backward.predictions, log.test$lprice))
print("Prediction Error Rate: ")
RMSE(log.backward.predictions, log.test$lprice)/mean(log.test$lprice)

log.backward.final.model <- lm(lprice~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance, data = log.nyc)
print("Full Model Adjusted R Squared: ")
print(summary(log.backward.final.model)$adj.r.squared)
print("Full Model AIC: ")
print(AIC(log.backward.final.model))
print("Full Model BIC: ")
print(BIC(log.backward.final.model))
print("Full Model RSS: ")
print(sum(resid(log.backward.final.model)^2))
summary(log.backward.final.model)
par(mfrow=c(2,2))
plot(log.backward.final.model)
vif(log.backward.final.model)
confint(log.backward.final.model)
```

SEQREP LINEAR
```{r seqrep selection model - linear}
linear.seqrep.select.model = regsubsets(price~.,data=linear.train,method="seqrep",nvmax=500)

par(mfrow=c(1,3))
srep.bic <- summary(linear.seqrep.select.model)$bic
plot(1:length(summary(linear.seqrep.select.model)$bic), srep.bic, type = "l", ylab = "BIC", xlab = "# Of Predictors")
srep.index.bic <- which(srep.bic == min(srep.bic))
points(srep.index.bic, srep.bic[srep.index.bic],col="red", pch=10)
print("")
print("Selected Number of Predictors From seqrep Selection (BIC)")
print(srep.index.bic)

srep.adjr2 <- summary(linear.seqrep.select.model)$adjr2
plot(1:length(summary(linear.seqrep.select.model)$adjr2), srep.adjr2, type = "l", ylab = "Adjusted R-squared", xlab = "# Of Predictors")
srep.index.adjr2 <- which(srep.adjr2 == max(srep.adjr2))
points(srep.index.adjr2, srep.adjr2[srep.index.adjr2],col="red", pch=10)
print("")
print("Selected Number of Predictors From seqrep Selection (AR2)")
print(srep.index.adjr2)

srep.rss <- summary(linear.seqrep.select.model)$rss
plot(1:length(summary(linear.seqrep.select.model)$rss), srep.rss, type = "l", ylab = "RSS", xlab = "# Of Predictors")
srep.index.rss <- which(srep.rss == min(srep.rss))
points(srep.index.rss, srep.rss[srep.index.rss],col="red", pch=10)
print("")
print("Selected Number of Predictors From seqrep Selection (RSS)")
print(srep.index.rss)

linear.seqrep.select.model2 = regsubsets(price~.,data=linear.train,method="seqrep",nvmax=(srep.index.bic)+1)
coef(linear.seqrep.select.model2,srep.index.bic)

linear.seqrep.train.model <-lm(price~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance,data=linear.train)
linear.seqrep.predictions <- linear.seqrep.train.model %>% predict(linear.test)
print("")
print("Test Adjusted R Squared: ")
print(R2(linear.seqrep.predictions, linear.test$price))
print("Test Root Mean Square Error: ")
print(RMSE(linear.seqrep.predictions, linear.test$price))
print("Test Mean Absolute Error: ")
print(MAE(linear.seqrep.predictions, linear.test$price))
print("Prediction Error Rate: ")
RMSE(linear.seqrep.predictions, linear.test$price)/mean(linear.test$price)

linear.seqrep.final.model <- lm(price~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance, data = linear.nyc)
print("Full Model Adjusted R Squared: ")
print(summary(linear.seqrep.final.model)$adj.r.squared)
print("Full Model AIC: ")
print(AIC(linear.seqrep.final.model))
print("Full Model BIC: ")
print(BIC(linear.seqrep.final.model))
print("Full Model RSS: ")
print(sum(resid(linear.seqrep.final.model)^2))
summary(linear.seqrep.final.model)
par(mfrow=c(2,2))
plot(linear.seqrep.final.model)
vif(linear.seqrep.final.model)
confint(linear.seqrep.final.model)
```

seqrep LOG
```{r seqrep selection model - log}
set.seed(77)

log.seqrep.select.model = regsubsets(lprice~.,data=log.train,method="seqrep",nvmax=500)

par(mfrow=c(1,3))
srep.bic <- summary(log.seqrep.select.model)$bic
plot(1:length(summary(log.seqrep.select.model)$bic), srep.bic, type = "l", ylab = "BIC", xlab = "# Of Predictors")
srep.index.bic <- which(srep.bic == min(srep.bic))
points(srep.index.bic, srep.bic[srep.index.bic],col="red", pch=10)
print("")
print("Selected Number of Predictors From seqrep Selection (BIC)")
print(srep.index.bic)

srep.adjr2 <- summary(log.seqrep.select.model)$adjr2
plot(1:length(summary(log.seqrep.select.model)$adjr2), srep.adjr2, type = "l", ylab = "Adjusted R-squared", xlab = "# Of Predictors")
srep.index.adjr2 <- which(srep.adjr2 == max(srep.adjr2))
points(srep.index.adjr2, srep.adjr2[srep.index.adjr2],col="red", pch=10)
print("")
print("Selected Number of Predictors From seqrep Selection (AR2)")
print(srep.index.adjr2)

srep.rss <- summary(log.seqrep.select.model)$rss
plot(1:length(summary(log.seqrep.select.model)$rss), srep.rss, type = "l", ylab = "RSS", xlab = "# Of Predictors")
srep.index.rss <- which(srep.rss == min(srep.rss))
points(srep.index.rss, srep.rss[srep.index.rss],col="red", pch=10)
print("")
print("Selected Number of Predictors From seqrep Selection (RSS)")
print(srep.index.rss)

log.seqrep.select.model2 = regsubsets(lprice~.,data=log.train,method="seqrep",nvmax=(srep.index.bic)+1)
coef(log.seqrep.select.model2,srep.index.bic)

log.seqrep.train.model <-lm(lprice~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance,data=log.train)
log.seqrep.predictions <- log.seqrep.train.model %>% predict(log.test)
print("")
print("Test Adjusted R Squared: ")
print(R2(log.seqrep.predictions, log.test$lprice))
print("Test Root Mean Square Error: ")
print(RMSE(log.seqrep.predictions, log.test$lprice))
print("Test Mean Absolute Error: ")
print(MAE(log.seqrep.predictions, log.test$lprice))
print("Prediction Error Rate: ")
RMSE(log.seqrep.predictions, log.test$lprice)/mean(log.test$lprice)

log.seqrep.final.model <- lm(lprice~neighbourhood_group+number_of_reviews+minimum_nights+calculated_host_listings_count+room_type+tsquare_distance, data = log.nyc)
print("Full Model Adjusted R Squared: ")
print(summary(log.seqrep.final.model)$adj.r.squared)
print("Full Model AIC: ")
print(AIC(log.seqrep.final.model))
print("Full Model BIC: ")
print(BIC(log.seqrep.final.model))
print("Full Model RSS: ")
print(sum(resid(log.seqrep.final.model)^2))
summary(log.seqrep.final.model)
par(mfrow=c(2,2))
plot(log.seqrep.final.model)
vif(log.seqrep.final.model)
confint(log.seqrep.final.model)
```